---
title: "Supplementary material, Bridging macroecology and macroevolution in the radiation of sigmodontine rodents"
author: "-"
date: "March 2022"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
<!-- badges: end -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Appendix S2

In this document we report supplementary results (mostly sensitivity analysis) of our manuscript. The figures and tables presented here follow the same order and settings as the ones shown in the main text:  

1 - The table with GLM coefficients showing differences in the slope of SES disparity ~ SES MPD between empirical and simulated disparity. Similar to Table 1. 

2 -  The relationship between SES MPD (Mean Pairwise (phylogenetic) Distance between species) and SES disparity produced by the empirical data set of skull morphology, and three data sets generated by trait simulation (Brownian motion[BM], Early-Burst[EB], and Ornstein-Uhlenbeck [OU]). Similar to Fig. 2 in the main text.

3 - A panel of five maps of A) empirical morphological disparity, B) disparity compared to a null model, C) disparity compared to a BM evolutionary model, D) significance of the map B, E) significance of the map C. Similar to Fig. 3 in the main text.

4 - A panel comprising four maps, showing the disparity expected by i) a null model, ii) BM model, iii) EB model, iv) OU model. Similar to Fig. 4 in the main text. We show the correlation between these expectations as a separate table after the map.    

The values of parameters, estimated through macroevolutionary models (using 'fitContinuous' in univariate datasets, and 'mvgls' in multivariate datasets), are only presented here in the supporting information.

\newpage

#### The relationship between species richness and disparity


```{r, echo = F, fig.height=8,fig.width=8,fig.align= "center",fig.margin=F, message=F, warning=F,out.width="50%",fig.keep =c(3,5,7,9)}

source ("./R/Interpretation_Fig3_Size_413spp.R")

# -------------------------------------------
# observed
# disparity in function of richness
par(mfrow=c(2,2), mar = c(4, 4, .1, .1))
plot(rowSums(presab),
     RAO_OBS$Observado,
     pch = 19,cex.axis=1,cex.lab=1,
     col = rgb(1,0,0,alpha=0.1),
     ylab = "Observed disparity",
     xlab= "Species richness")
# model
mod<-lm(RAO_OBS$Observado ~ poly(rowSums(presab),2))
# predict
lines (rowSums(presab)[order(rowSums(presab))],
       predict(mod)[order(rowSums(presab))],
       lwd=2)

# the SES
plot(rowSums(presab),
      RAO_OBS$SES,
     pch = 19,cex.axis=1,cex.lab=1,
     col = rgb(0,0,1,alpha=0.1),
     ylab = "SES RAO\n(obsRAO - mean(rdmRAO)/sd(rdmRAO))",
     xlab = "Species richness")
# model
mod2<-lm(RAO_OBS$SES ~ poly(rowSums(presab),2))
# predict
lines (rowSums(presab)[order(rowSums(presab))],
       predict(mod2)[order(rowSums(presab))],
       lwd=2)

# SES OU
plot(rowSums(presab),
     rowMeans(do.call(cbind,sapply(RAO_OU, "[","Observado",simplify=T))),
     pch = 19,cex.axis=1,cex.lab=1,
     col = rgb(1,0,0,alpha=0.1),
     ylab = "OU disparity",
     xlab= "Species richness")
# model
mod<-lm(rowMeans(do.call(cbind,sapply(RAO_OU, "[","Observado",simplify=T))) ~ poly(rowSums(presab),2))
# predict
lines (rowSums(presab)[order(rowSums(presab))],
       predict(mod)[order(rowSums(presab))],
       lwd=2)

# the SES
plot(rowSums(presab),
      SES_NEUTRAL,
     pch = 19,cex.axis=1,cex.lab=1,
     col = rgb(0,0,1,alpha=0.1),
     ylab = "SES RAO OU\n(obsRAO - mean(RAO OU)/sd(RAO OU))",
     xlab = "Species richness")
# model
mod2<-lm(SES_NEUTRAL ~ poly(rowSums(presab),2))
# predict
lines (rowSums(presab)[order(rowSums(presab))],
       predict(mod2)[order(rowSums(presab))],
       lwd=2)




```

Fig. S2.1. Relationship between assemblage-level species richness (SR) and observed/empirical values of phenotypic disparity (Rao's entropy index) (top left) and OU disparity (bottom left). In the right portion of this panel (plot with blue points), we show the relationship between SR and disparity after comparing observed disparity to a randomization-based null model (top right) and an OU model of evolution (bottom right). 


\newpage

#### Results, considering the complete set of species, and using the complete sample of 100 fully resolved phylogenies from Upham et al. (2019) (i.e., considering phylogenetic uncertainty). The following figures are supplementary to Figs. 2 to 5, and tables 1 and 2 presented in the main text.


```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F,fig.keep ='last'}

source ("./R/Interpretation_Fig2_and_statistics_Size_413spp.R")

plot(params_estim)

```

Fig. S2.2. Density plot showing the estimates of each model parameters for univariate and multivariate models. In the top we show the estimates of the parameter 'sigma' for the BM, EB, and OU models of evolution (evolutionary models for univariate trait data (fitContinuous)). In the middle, we show the parameters 'beta' and 'alpha', which are specific from EB and OU, respectively (evolutionary models for univariate trait data (fitContinuous)). In the bottom, we show estimates of sigma for the BM model (evolutionary models for multivariate trait data (mvGLS)). In univariate models, the density represents variation in estimates across the 100 fully resolved phylogenies. In the multivariate model, the density represents variation of sigma across traits and phylogenies (112 simulated ‘landmarks’ and 100 fully resolved phylogenies).Results produced by using the dataset of 413 species.

<!-- badges: start -->
<!-- badges: end -->


Below we present a table of GLM coefficients that considers phylogenetic uncertainty when testing which model produced a simulated disparity closer to the empirical disparity (results alternative to Table 1 in the main text) show that the disparity simulated by the OU model gets closer to the empirical disparity (Table S2.1). The GLM estimates were averaged (and its standard deviation calculated) across 100 runs per phylogeny (for 100 fully resolved phylogenies).


Table S2.1: Averaged GLM estimates, obtained by averaging estimates produced by each one of the 100 different phylogenies used to simulate traits and calculate disparity. Results produced by using the dataset of 413 species.

```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.margin=F, message=F, warning=F,fig.show="hide"}

 (list.mean)

```
Table S2.2: Standard deviation of GLM estimates, obtained by calculating the standard deviation of estimates produced by each one of the 100 different phylogenies used to simulate traits and calculate disparity. Results produced by using the dataset of 413 species.
```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.margin=F, message=F, warning=F,fig.show="hide"}

list.sd

```

\newpage

### Analysis with 216 species, and considering phylogenetic uncertainty

Using this data subset, the estimated parameters of macroevolutionary models were very similar to those estimated for the complete set of species (compare Figs. S2.1 and S2.2). we then produced a bivariate plot describing the relationship between SES disparity and SES MPD, for the empirical and simulated data sets. We found that this relationship (Fig. S2.3) was similar to the one reported in the main results (Fig. 2). Overall, the disparity produced by the OU model had a closer relationship with the empirical disparity than had the other models of evolution (BM and EB) (Fig. S2.3).


```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F,fig.keep ='last'}

source ("./R_uncertainty_S2/Interpretation_Fig2_and_statistics_Size_216spp.R")

plot(panel_params)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S2.3.Density plot showing the estimates of each model parameters for univariate and multivariate models. In the top we show the estimates of the parameter 'sigma' for the BM, EB, and OU models of evolution (evolutionary models for univariate trait data (fitContinuous)). In the middle, we show the parameters 'beta' and 'alpha', which are specific from EB and OU, respectively (evolutionary models for univariate trait data (fitContinuous)). In the bottom, we show estimates of sigma for the BM model (evolutionary models for multivariate trait data (mvGLS)). In univariate models, the density represents variation in estimates across the 100 fully resolved phylogenies. In the multivariate model, the density represents variation of sigma across traits and phylogenies (112 simulated ‘landmarks’ and 100 fully resolved phylogenies).Results produced by using the dataset of 216  species.

<!-- badges: start -->
<!-- badges: end -->

```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F,fig.keep = 2}


plot(panel)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S2.4. Bivariate plot showing the relationship between SES MPD and SES disparity, for empirical and simulated disparity. Results produced using the data set of 216 species.

<!-- badges: start -->
<!-- badges: end -->

As presented in the results (mapping section), we counted the number of cells with values of disparity higher, equal, or lower than the null model (randomization-based null model, NULL) and the evolutionary models. We found that 364 communities had a disparity lower than expected by the EB and OU models, and 344 communities had a disparity lower than the BM model. Nonetheless, only three communities had disparity lower than the null disparity. On the other hand, 182 communities had disparity higher than the expected by the OU model, 138, 129, and 90 communities had disparity higher than the expected by the BM, EB and NULL models.  



```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.margin=F, message=F, warning=F,fig.show="hide"}


count_cells_relative_to_models


```

<!-- badges: start -->
<!-- badges: end -->

When mapping the values of empirical and simulated trait values, we found a high agreement between the complete dataset and the dataset with 216 species (Fig. S2.4). Overall, we found a prevalence of higher disparity than expected by an OU model in two portions of the Atlantic Rainforest, and lower disparity than expected by an OU model in Amazon, Northeastern Brazil, and south Andes.


```{r, echo = F, fig.height=10,fig.width=5,fig.margin=T,fig.align="center", message=F, warning=F}

source ("./R_uncertainty_S2/Interpretation_Fig3_Size_216spp.R")

# plot (panel2)
# run this to get the panel with the relationship between MPD and DISPARITY
# and other results


```
Fig. S2.5. Map of empirical (A), null (random shuffling of species in trait matrix) (B) and simulated disparity (using the OU model) (C). In D we show the significance of SES values presented in B, and in E we show the significance of SES values presented in C. Results produced by using the dataset of 216 species.  
  
<!-- badges: start -->
<!-- badges: end -->

Finally, we show the maps of null and simulated disparity (Fig. S2.5). We found that using a null model is quite similar to simulating a trait using the OU model. This can be seen in the following map (and the correlations presented after the Fig. S2.5), where the null disparity and the disparity produced by the OU model produced quite similar, highly correlated maps (0.96, Table below the map). The correlation was also high  between  Null and BM disparity.


```{r, echo = F, fig.height=7,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F}


alternative_map1


```
Fig. S2.6. Map of null (random shuffling of species in trait matrix) and simulated disparity values (using the BM, OU and EB models of evolution). The legend is common to all maps. Results produced by using the dataset of 216 species.  


Table of correlation between these estimates


```{r , echo=FALSE,fig.margin=F, message=F, warning=F}

## show the correlation between average null and simulated by OU

cor (data.frame (RAO_OBS$med_nulo, obsBM,obsEB,obsOU))


```


\newpage


### Results for subclade Oryzomyalia, considering phylogenetic uncertainty

Bivariate plot similar to Fig. 2. This is very similar to the plots reported in the main text and sensitivity analysis with 216 species (just presented). However, the main difference is that we lost values of SES MPD higher than zero, as now we lost phylogenetic distances by using only data from the subclade Oryzomyalia.



```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F,fig.keep ='last'}

source ("./R_uncertainty_S2/Interpretation_Fig2_and_statistics_Size_oryzomyialia.R")

 plot(panel_params)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```

Fig. S2.7. Density plot showing the estimates of each model parameters for univariate and multivariate models. In the top we show the estimates of the parameter 'sigma' for the BM, EB, and OU models of evolution (evolutionary models for univariate trait data (fitContinuous)). In the middle, we show the parameters 'beta' and 'alpha', which are specific from EB and OU, respectively (evolutionary models for univariate trait data (fitContinuous)). In the bottom, we show estimates of sigma for the BM model (evolutionary models for multivariate trait data (mvGLS)). In univariate models, the density represents variation in estimates across the 100 fully resolved phylogenies. In the multivariate model, the density represents variation of sigma across traits and phylogenies (112 simulated ‘landmarks’ and 100 fully resolved phylogenies). Results produced by using the dataset of Oryzomyalia species.

<!-- badges: start -->
<!-- badges: end -->

```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F,fig.keep =2}

plot (panel)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```

Fig. S2.8. Bivariate plot showing the relationship between SES MPD and SES disparity, for empirical and simulated disparity. Results produced by using the dataset of Oryzomyalia species.

<!-- badges: start -->
<!-- badges: end -->

Counting the number of cells. We found that 269 communities had a disparity lower than expected by the OU model, 262 had a disparity lower than expected by the EB model, and 247 communities had a disparity lower than the BM model. Nonetheless, no community had disparity lower than the null disparity. On the other hand, 109 communities had disparity higher than the expected by the NULL model, and only 19, 18 and 16 communities had disparity higher than the expected by the the OU, BM and EB models of evolution.  

```{r , echo=FALSE,fig.margin=F, message=F, warning=F}


count_cells_relative_to_models


```

<!-- badges: start -->
<!-- badges: end -->


The maps of observed and simulated disparity still resemble the ones shown in Fig. 3 (main text), and the ones just reported considering 216 species (Fig. S2.4). However, less communities had significantly lower and higher disparity than the disparity expected by the OU model (Fig. S2.9).



```{r, echo = F, fig.height=10,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}

source ("./R_uncertainty_S2/Interpretation_Fig3_Size_oryzomyialia.R")

# plot(panel2)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S2.9. Map of empirical (A), null (random shuffling of species in trait matrix) (B) and simulated disparity (using the OU model) (C). In D we show the significance of SES values presented in B, and in E we show the significance of SES values presented in C. Results produced by using the dataset of Oryzomyalia species.  

<!-- badges: start -->
<!-- badges: end -->

Finally, we report high correlation between null and simulated disparity (map similar to Fig. 4, and Fig. S2.3 presented above).



```{r, echo = F, fig.height=7,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F}


alternative_map1


```
Fig. S2.10. Map of null (random shuffling of species in trait matrix) and simulated disparity values (using the BM, OU and EB models of evolution). The legend is common to all maps. Results produced by using the dataset of Oryzomyalia species.  



Correlations between maps were similar to the ones reported in the main text and in the previous analyses (the highest correlation was found between NULL and OU-simulated disparity.



```{r , echo=FALSE,fig.margin=F, message=F, warning=F}

## show the correlation between average null and simulated by OU

cor (data.frame (RAO_OBS$med_nulo, obsBM,obsEB,obsOU))


```

\newpage

### Appendix S3 

#### Results considering the consensus phylogeny. As it has 285 tips, the number of species used in trait simulations was 285 (for trait simulated with all tips), being 169 with occurrence, trait and phylogenetic data.  


The figures we will present now have the same settings and order as the ones we presented above.

Parameter estimates for univariate (below) and multivariate trait simulations (Fig. S3.1).

```{r, echo = F, fig.height=4,fig.width=4,fig.align= "center",fig.margin=F, message=F, warning=F}

source ("./R_concensus_supp_S3/Interpretation_Fig2_and_statistics_Size_CONS.R")

# run this to get the panel with the relationship between MPD and DISPARITY
(df_density)


```



```{r, echo = F, fig.height=4,fig.width=4,fig.align= "center",fig.margin=F, message=F, warning=F}

# run this to get the panel with the relationship between MPD and DISPARITY
plot(fig_params_multiv)


```
Fig. S3.1: Density plot of the sigma parameter, as estimated by the BM macroevolutionary model fitted to skull shape data. Results produced by using the data set of 285 species in the consensus tree.

<!-- badges: start -->
<!-- badges: end -->

Bivariate plot similar to Fig. 2. The OU-resulting disparity is still closer to the empirical disparity than other models (see the table of coefficients and difference).  


```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F}

# run this to get the panel with the relationship between MPD and DISPARITY
plot(figure1)


```

Fig. S3.2. Bivariate plot showing the relationship between SES MPD and SES disparity, for empirical and simulated disparity. Results produced by using the data set of 285 species in the consensus tree.

<!-- badges: start -->
<!-- badges: end -->

```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F}
  
(av_coeff)

```

Counting the number of cells. We found a large number of cells with disparity lower than the BM and EB models, and a small number with disparity lower than NULL and OU models. We found that 1271 communities had disparity higher than expected by the OU model, and much less communities with disparity higher than BM, EB and NULL models.  



```{r , echo=FALSE,fig.margin=F, message=F, warning=F}


count_cells_relative_to_models


```

<!-- badges: start -->
<!-- badges: end -->

Maps similar to Fig. 3, and to maps just reported considering the subsets of the data (Fig. S2.2 and S2.5). The maps still resemble, but now we did not find cells with disparity higher or lower than expected by BM model.



```{r, echo = F, fig.height=10,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}

source ("./R_concensus_supp_S3/Interpretation_Fig3_Size_CONS.R")

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

# plot(panel2)

```
Fig. S3.3. Map of empirical (A), null (random shuffling of species in trait matrix) (B) and simulated disparity (using the BM model) (C). In D we show the significance of SES values presented in B, and in E we show the significance of SES values presented in C. Results produced by using the data set of 285 species in the consensus tree.       

  
Using the consensus phylogeny, we found variation relative to results presented above and in the main text. The maps of BM- and OU-simulated disparity showed higher spatial variation than reported previously. 


```{r, echo = F, fig.height=7,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F}


alternative_map1


```
Fig. S3.4. Map of null (random shuffling of species in trait matrix) and simulated disparity values (using the BM, OU and EB models of evolution). The legend is common to all maps. Results produced by using the data set of 285 species in the consensus tree.    
  
<!-- badges: start -->
<!-- badges: end -->

The correlation between null disparity and simulated disparity was moderate to low. The highest correlation we found was between OU- and EB-simulated disparity (0.38). The correlation between null disparity and OU disparity was 0.19.



```{r , echo=FALSE,fig.margin=F, message=F, warning=F}

## show the correlation between average null and simulated by OU

cor (data.frame (RAO_OBS$med_nulo, obsBM,obsEB,obsOU))


```

\newpage

## Appendix S4 

#### Results considering skull shape, and the application of macroevolutionary models for multivariate trait simulations (using a BM model of evolution). The  estimates of model parameters for datasetes of 216 species and Oryzomyalia species were presented in Figs. S2.2 and S2.6. 

None community had disparity higher than either NULL or BM models. Nonetheless, most communities had disparity lower than the null model.

```{r , echo=FALSE, fig.show="hide",fig.margin=F, message=F, warning=F}
source ("./R_uncertainty_S2/Interpretation_Fig2_and_statistics_Shape_216spp.R")


count_cells_relative_to_models


```

<!-- badges: start -->
<!-- badges: end -->

The BM model had a good fit to skull shape (Fig. S3.1, Table S3.1).


```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F,fig.keep =2}


plot(panel)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S4.1. Bivariate plot showing the relationship between SES MPD and SES disparity, for empirical and simulated disparity (multivariate trait simulations using Brownian motion model evolution). Results produced by using the dataset of 216 species. 

<!-- badges: start -->
<!-- badges: end -->

Table S4.1: Averaged GLM estimates, obtained by averaging estimates produced by each one of the 100 different phylogenies used to simulate traits and calculate disparity. Results produced by using the dataset of 216 species.

```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.margin=F, message=F, warning=F,fig.show="hide"}


 (list.mean)

```


Table S4.2: Standard deviation of GLM estimates, obtained by calculating the standard deviation of estimates produced by each one of the 100 different phylogenies used to simulate traits and calculate disparity. Results produced by the data set of 216 species.
```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.margin=F, message=F, warning=F,fig.show="hide"}

list.sd

```

<!-- badges: start -->
<!-- badges: end -->

Assemblages presenting lower disparity than predicted by a BM model of evolution were located at the Chaco, Amazon Basin, north Brazil, South Andes and Patagonia (Fig. S3.2).


```{r, echo = F, fig.height=10,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}

source ("./R_uncertainty_S2/Interpretation_Fig3_Shape_216spp.R")

# plot(panel2)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S4.2. Map of empirical (A), null (random shuffling of species in trait matrix) (B) and simulated disparity (using the BM model) (C). In D we show the significance of SES values presented in B, and in E we show the significance of SES values presented in C. Results produced by using the dataset of 216 species. 

<!-- badges: start -->
<!-- badges: end -->

```{r, echo = F, fig.height=3,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}


plot(alternative_map1)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S4.3. Map of null (random shuffling of species in trait matrix) and simulated disparity values (using the BM model of evolution to simulate multivariate trait datasets). The legend is common to all maps. Results produced by using the dataset of 216 species.  


```{r, echo = F, fig.height=5,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}


cor (data.frame (RAO_OBS$med_nulo, 
                 obsBM))

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```

### Results for subclade Oryzomyalia, considering phylogenetic uncertainty


```{r , echo=FALSE,fig.margin=F, message=F, warning=F}
source ("./R_uncertainty_S2/Interpretation_Fig2_and_statistics_Shape_oryzomyialia.R")


count_cells_relative_to_models


```

Again, the BM model had a good fit to skull shape (Fig. S3.1, Table S3.1).


```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F}


plot(panel)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S4.4. Bivariate plot showing the relationship between SES MPD and SES disparity, for empirical and simulated disparity (multivariate trait simulations using Brownian motion model evolution). Results produced by using the dataset of Oryzomyalia species. 

<!-- badges: start -->
<!-- badges: end -->

Table S4.3: Averaged GLM estimates, obtained by averaging estimates produced by each one of the 100 different phylogenies used to simulate traits and calculate disparity. Results produced by using the dataset of Oryzomyalia species. 

```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.margin=F, message=F, warning=F,fig.show="hide"}


 (list.mean)

```

Table S4.4: Standard deviation of GLM estimates, obtained by calculating the standard deviation of estimates produced by each one of the 100 different phylogenies used to simulate traits and calculate disparity. Results produced by using the dataset of Oryzomyalia species. 
```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.margin=F, message=F, warning=F,fig.show="hide"}

list.sd

```

```{r, echo = F, fig.height=10,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}

source ("./R_uncertainty_S2/Interpretation_Fig3_Shape_oryzomyialia.R")

# plot(panel2)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S4.5. Map of empirical (A), null (random shuffling of species in trait matrix) (B) and simulated disparity (using the BM model) (C). In D we show the significance of SES values presented in B, and in E we show the significance of SES values presented in C. Results produced by using the dataset of Oryzomyalia species. 

```{r, echo = F, fig.height=3,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}


plot(alternative_map1)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S4.6. Map of null (random shuffling of species in trait matrix) and simulated disparity values (using the BM model of evolution to simulate multivariate trait datasets). The legend is common to all maps. Results produced by using the dataset of Oryzomyalia species. 

<!-- badges: start -->
<!-- badges: end -->


```{r, echo = F, fig.height=5,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}


cor (data.frame (RAO_OBS$med_nulo, 
                 obsBM))

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```

## Results produced by using the dataset of species in the consensus tree.

```{r pressure, echo=FALSE,fig.margin=F, message=F, warning=F}
source ("./R_concensus_supp_S3/Interpretation_Fig2_and_statistics_Shape_CONS.R")


count_cells_relative_to_models


```

<!-- badges: start -->
<!-- badges: end -->

Again, the BM model had a good fit to skull shape (Fig. S3.1, Table S3.1).


```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.align= "center",fig.margin=F, message=F, warning=F}


plot(figure1)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S4.7. Bivariate plot showing the relationship between SES MPD and SES disparity, for empirical and simulated disparity (multivariate trait simulations using Brownian motion model evolution). Results produced by using the dataset of species in the consensus tree.

<!-- badges: start -->
<!-- badges: end -->

Table S4.5: Averaged GLM estimates, obtained by averaging estimates produced by each one of the 100 different phylogenies used to simulate traits and calculate disparity. Results produced by using the dataset of species in the consensus tree.

```{r, echo = F, fig.height=6.5,fig.width=6.5,fig.margin=F, message=F, warning=F,fig.show="hide"}


 summary(av_coeff)


```


```{r, echo = F, fig.height=10,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}

source ("./R_concensus_supp_S3/Interpretation_Fig3_Shape_CONS.R")

# plot(panel2)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S4.8. Map of empirical (A), null (random shuffling of species in trait matrix) (B) and simulated disparity (using the BM model) (C). In D we show the significance of SES values presented in B, and in E we show the significance of SES values presented in C. Results produced by using the dataset of species in the consensus tree. 

<!-- badges: start -->
<!-- badges: end -->

```{r, echo = F, fig.height=3,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}


plot(alternative_map1)

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S4.9. Map of null (random shuffling of species in trait matrix) and simulated disparity values (using the BM model of evolution to simulate multivariate trait datasets). The legend is common to all maps. Results produced by using the dataset of species in the consensus tree. 



```{r, echo = F, fig.height=10,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}


cor (data.frame (RAO_OBS$med_nulo, 
                 obsBM))

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```



```{r, echo = F, fig.height=3,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F, fig.keep ="last"}

source ("./R/Interpretation_Fig3_Shape_413spp.R")

alternative_map1
# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```
Fig. S4.10. Map of null (random shuffling of species in trait matrix) and simulated disparity values (using the BM model of evolution to simulate multivariate trait datasets). The legend is common to all maps. Results produced by the dataset of 413 species and the set of 100 phylogenies. 



```{r, echo = F, fig.height=10,fig.width=5,fig.align= "center",fig.margin=T,fig.align="center",  message=F, warning=F}


cor (data.frame (RAO_OBS$med_nulo, 
                 obsBM))

# run this to get the panel with the relationship between MPD and DISPARITY
# and other results

```